---
title: "summarize-layer-by-huc12"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{summarize-layer-by-huc12}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteEval{FALSE}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(middlesnake)
library(dplyr)
library(sf)
```


# Get watersheds by Conservation District
automatically clips watersheds to district boundary
```{r}

columbia_huc12 <- get_huc_by_cd("Columbia")

```

# Show geoserver layers

```{r}

show_geoserver_layers()

```


# Get the layer that need summarized 

Here we use frequently flooded areas
```{r}

flooded <- get_geoserver_layer("columbia-frequently-flooded")

```

# Summarize

How many frequently flooded acres are inside each huc 12 watershed in Columbia County?
```{r}

columbia_flooded_huc12 <- summarize_acres_by_huc12(huc12 = columbia_huc12,
                                                   overlay = flooded)
```

returns `huc12` (numeric code)  and `acres`. 
If you want the watershed name add it
```{r}
columbia_flooded_huc12 %>%
    dplyr::left_join(columbia_huc12 %>% 
                       select(name,huc12)
                     )
```

# Let's try wetlands

```{r}

wetlands <- get_geoserver_layer("columbia-wetlands")

```


# This time we want to stratify by wetland type within each huc 12
We'll set the `group_vars` argument
```{r}

columbia_wetlands_huc12 <- summarize_acres_by_huc12(huc12 = columbia_huc12,overlay = wetlands,group_vars = "WETLAND_TY")

columbia_wetlands_huc12

```

# Use a layer not on Geoserver

We already have Columbia watersheds to use as aoi (i.e. Area of Interest)
to automatically subset the layer we are grabbing. In this example is using WA DNR Priority areas.
```{r}
dnr_forest_priority <- get_gis_from_url(url = "https://services.arcgis.com/4x406oNViizbGo13/arcgis/rest/services/Eastern_WA_Forest_Health_Priority_HUC6_Watersheds/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=geojson",aoi = columbia_huc12)
```

```{r}


columbia_forest_priority_huc12 <- summarize_acres_by_huc12(huc12 = columbia_huc12,
                                                            overlay = dnr_forest_priority)

```

